---
title: "SCC and Policy Choice Under Uncertainty"
author: "Bri Baker, Mikel Irigoyen"
date: "5/23/2021"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    highlight: kate
  pdf_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(decisionSupport)

```

## Data Setup

```{r data}

war <- read_csv(here("warming.csv")) %>%  # add warming data
  rename("id" = "X1")

dam <- read_csv(here("damages.csv")) # add damages data

pulse <- 35000000000 # tons of carbon

```

## Question 1

```{r damage function}
dam <- dam %>% 
  mutate(warm2 = warming^2) # make quadratic term

dam_deg <- lm(damages ~ 0 + warming + warm2, data = dam) # fit quadratic

dam_deg_a <- dam_deg$coefficients[[2]] # squared term
dam_deg_b <- dam_deg$coefficients[[1]] # second term
  
ggplot(dam, aes(x = warming, y = damages)) + # plot
  geom_point() + # scatter
  geom_smooth()+
  labs(title="Damage Function",
        x ="Change in Global Mean Temp", y = "Damages ($)")+
  theme_minimal()

```

Damage = 1.94e+13 x warming^2^ + -1.03e+12 x warming.


## Question 2

```{r graphs}

war <- war %>% 
  mutate(damages_baseline = dam_deg_a*warming_baseline^2 + dam_deg_b*warming_baseline, # calc damages baseline
         damages_pulse = dam_deg_a*warming_pulse^2 + dam_deg_b*warming_pulse, # calc damages with pulse
         diff_damages = damages_pulse-damages_baseline, # calc difference in damages
         diff_ton = diff_damages/pulse # calc difference per ton
         )

ggplot(war, aes(x = year, y = damages_baseline))+
  geom_point() +
  labs(title="Damages Without the Pulse",
        x ="Year", y = "Damages ($)")+
  theme_minimal() 

ggplot(war, aes(x = year, y = damages_pulse))+
  geom_point() +
  labs(title="Damages With the Pulse",
        x ="Year", y = "Damages ($)")+
  theme_minimal() 

ggplot(war, aes(x = year, y = diff_damages))+
  geom_point() +
  labs(title="Difference in Damages With and Without Pulse",
        x ="Year", y = "Damages ($)")+
  theme_minimal() 

ggplot(war, aes(x = year, y = diff_ton))+
  geom_point() +
  labs(title="Difference in Damages Over Time From the Pulse Per Ton of
CO2 ",
        x ="Year", y = "Damages ($)")+
  theme_minimal() 

```


## Question 3

```{r scc}

discounts <- c(.01, .02, .03, .04, .05) # set range of discounts
scc <- tribble(~c1, ~c2)  #initiate df
for (i in discounts){
  disc <- i # assign discount
  scc_pulse <- discount(war$diff_ton, # calc npv of differences using sum and discounting
                  disc,
                  calculate_NPV = TRUE)
  bri <- c(disc, scc_pulse) # make list of discount rate and pulse
  scc <- rbind(scc, bri) # add to df
} 
scc <- scc %>% 
  rename("Discount_Rate" = "X0.01",
         "SCC" = "X183.316049160008") # name df columns for clarity

ggplot(scc)+ # plot
  geom_line(aes(x = Discount_Rate, y = SCC))+
  labs(title="Social Cost of Carbon",
        x ="Discount Rate (%)", y = "SCC (Dollars)")+
  theme_minimal()
```


## Question 4

```{r}
#r = p + ng
r = 0.001 +(2)*(0.01) # calculate r using ramsey rule

scc_model <- lm(SCC ~ Discount_Rate, data = scc)# calc scc line based on range

slope <- scc_model$coefficients[[2]] # pull slope
intercept <- scc_model$coefficients[[1]] #pull intercept

scc_science <- r*slope + intercept # calculate scc with ramsey rule r

ggplot(scc)+ # plot
  geom_line(aes(x = Discount_Rate, y = SCC))+
  geom_point(x = r, y = scc_science)+
  labs(title="SCC Using Ramsey Rule",
        x ="Discount Rate (%)", y = "SCC (Dollars)")+
  theme_minimal()

```


The SCC is $`r round(scc_science, 2)`/ton.


## Question 5

```{r policy a and policy b}

war <- war %>% 
  mutate(warming_a = warming_baseline*1.5, # warming in a1
         damages_a = dam_deg_a*warming_a^2 + dam_deg_b*warming_a, # damages in a 1
         warming_b = case_when(year >= 2050 ~ 1.29, # warming in b
                               TRUE ~ warming_baseline),
         damages_b = dam_deg_a*warming_b^2 + dam_deg_b*warming_b) # damages in b

npv_a1 <- discount(war$damages_baseline, # npv of damages of a1
                   discount_rate = 0.02,
                   calculate_NPV = TRUE)

npv_a2 <- discount(war$damages_a, # npv of damages of a2
                   discount_rate = 0.02,
                   calculate_NPV = TRUE)

npv_b <- discount(war$damages_b, # npv of damages of b
                  discount_rate = 0.02,
                  calculate_NPV = TRUE)

npv_a <- (npv_a2 - npv_a1)/2 + npv_a1 # calc "utility" of a

cost_difference <- npv_a - npv_b # compare costs of a and b
  
```

The expected present value of damages up to 2100 under Policy A is either $1.28e+16 or $5.66e+15.
The expected present value of damages up to 2100 under Policy B is $1.89e+15.

Until X is $7.36e+15 it still makes economic sense to pursue Policy b instead of Policy A. 

If society were risk averse, the utility function would be concave rather than linear which would reduce the utility of the uncertain outcome.
